Z-Resonance at LEP

### 2010-08-04 Z0-Publikation/tex/paper-template.tex
roberts hadr. breite korrigiert, alles ein bisschen besser

\title{Z-Resonance at LEP}
\titlerunning{Z-Resonance at LEP}
\authorrunning{Thomas Murach, Robert Riemann}
\author{{\bf Draft Version 1.1}\\
\medskip \\
Thomas Murach,$^{1}$
Robert Riemann$^{1}$
} 
\institute{$^1$Department of Physics, Humboldt University of Berlin, Germany\\}  
\date{Received: \today / Revised version: 0.1}
\abstract{
This paper describes the analysis of the L3 detector data taken at different
center of mass energies around \SI{90}{\GeV}, which reveals the properties of
the $Z$-Boson. Additionally the weak mixing angle and the colour factor are determined.
The Z-boson mass is evaluated to $M_Z = \SI{91.16(3)}{\GeV}$ with a decay width
of $Γ_Z = \SI{2.61(6)}{\GeV}$.
The weak mixing angle is computed to $\sin² Θ_W = \num{0.23(9)}$ and the colour
factor to $N_C = \num{3.2(13)}$.
All resuls are in good agreement to the standard model and can be considered as
a confirmation of it.
}

@@

Due to inefficiencies in energy measurement it is possible that not all the
energy provided by the apparatus can finally be seen. Therefore there are large
uncertainties in jet energy measurements, whereas these uncertainties are
smaller for muonic events. This fact can also be used to distinguish these
events from each other and from background as well. For example if the $Z^0$
decays to two very unstable \todo{plural von τ?}\ tauons, these tauons decay very fast. Then two
neutrinos are created, but they cannot be measured, so that their contribution
to the visible energy is missing. The visible part of the tau decays resembles
hadronic events, so it is important to be able to separate this background from
the hadronic signal.

@@

\subsection{Determination of the leptonic decay width}
Through the knowledge of the overall cross section it is possible to calculate
the leptonic decay width $Γ_l$. First of all we assume that the decay width is
the same for all leptons. We can use the following formula for determining it
by using the fit parameter for the muonic cross section:
\begin{equation}
σ_0^μ = \frac{12π}{M_Z^2}\frac{Γ_l^2}{Γ_Z^2}.
\label{eqn:leptonicwidth}
\end{equation}
This equation can easily be solved to determine the resulting leptonic decay
width $Γ_l$, because all the other variables are known.
\begin{equation}
Γ_l = (91 \pm 2 \pm 1)\si{\mega\electronvolt}
\label{eqn:leptonicwidthresult}
\end{equation}
In order to calculate the leptonic decay width, it is also possible to utilize
the following equation:
\begin{equation}
Γ_l = \frac{1}{3}(Γ_Z - 3Γ_{\mathrm{Had}} - 3Γ_{ν_l}).
\end{equation}
$Γ_{\mathrm{Had}}$ can be calculated in the same way as the leptonic decay
width in \eref{leptonicwidth}. $Γ_{ν_l}$ can be determinde by using the
following relation:
\begin{eqnarray}
Γ_{ν_e} &=& \frac{G_FM_Z^3}{12\sqrt{2}π}\\
&=& (165.71 \pm 0.10 \pm 0.09)\si{\mega\electronvolt}
\end{eqnarray}
Now the leptonic decay width $Γ_l$ can be evaluated.
\begin{equation}
Γ_l = (83 \pm 4 \pm 3) \si{\mega\electronvolt}
\end{equation}
This is in the same order of magnitude as the result shown in
\eref{leptonicwidthresult}.

@@

The achieved results by now offer furthermore an oportunity to obtain the
weak mixing angle using the formula
\begin{eqnarray}
Γ_l &=& \frac{G_F M_Z^3}{24 \sqrt{2}π}\left(1+[1-4|Q_l|\sin² Θ_W]²\right)\,.
\end{eqnarray}
As $Γ_l$ as well as $M_Z$ is known, the expression $\sin² Θ_W$ can be computed.
\begin{eqnarray}
\sin^2\Theta_W &=& \frac{1-\sqrt{\frac{\Gamma_e\cdot 24\sqrt{2}\pi}{G_FM_Z^3}-1}}{4}\\
\nonumber &=& 0.23 \pm 0.05 \pm 0.08
\end{eqnarray}
Thereby the weak mixing angle is determined, too.
\begin{eqnarray}
Θ_W &=& \arcsin(\sqrt{\sin^2\Theta_W})\\
\nonumber &=& 0.50 \pm 0.11 \pm 0.07\\
\nonumber &=& (28 \pm 6 \pm 4)\symbol{23}
\end{eqnarray}

@@

Finally the measurements allow also to calculate the colour factor $N_C$ using
$Γ_{\mathrm{had}}$. The hadronic decay width can be obtained by two formulas.
This width can be computed, as the overall $Γ_Z$ in known, but also the components
$Γ_{ν_e}$ and $Γ_l$.
\begin{equation}
 Γ_Z = Γ_{\mathrm{had}} + 3Γ_{ν_e} + 3Γ_l
 \label{eqn:gammahad1}
\end{equation}
In reference to the standard model there is a second way to compute $Γ_{\mathrm{had}}$.
\begin{equation}
 Γ_{\mathrm{had}} = N_C \cdot K_{QCD}\cdot ( 2Γ_u + 3Γ_d )
 \label{eqn:gammahad2}
\end{equation}
Replacing $Γ_{\mathrm{had}}$ in \eref{gammahad2} using \eref{gammahad1} leads to a
formula which allows to compute the colour factor $N_C$.
\begin{equation}
 N_C = 3.20 \pm 0.11 \pm 0.07
\end{equation}
This result confirms, that there are no new quarks to be expected with a mass
lower than some \SI{90}{\GeV}.

\todo[inline,size=]{
kleine erläuterung zur abweichung, kleines schlusswort
in zeile 304 steht schon, wie man die hadronische breite berechnen könnte. das
haben wir im protokoll nicht gemacht, sondern nur ne andere methode. diese
methode mit reinnehmen? 
}

@@

\section{ Summary}

>>>>

\title{Z-Resonance at LEP}
\titlerunning{Z-Resonance at LEP}
\authorrunning{Thomas Murach, Robert Riemann}
\author{{\bf Draft Version 1.1}\\
\medskip \\
Thomas Murach,$^{1}$
Robert Riemann$^{1}$
} 
\institute{$^1$Department of Physics, Humboldt University of Berlin, Germany\\}  
\date{Received: \today / Revised version: 0.1}
\abstract{
This paper describes the analysis of the L3 detector data taken at different
center of mass energies around \SI{90}{\GeV}, which reveals the properties of
the $Z$-Boson. Additionally the weak mixing angle and the colour factor are determined.
The Z-boson mass is evaluated to $M_Z = \SI{91.16(3)}{\GeV}$ with a decay width
of $Γ_Z = \SI{2.61(6)}{\GeV}$.
The weak mixing angle is computed to $\sin² Θ_W = \num{0.23(9)}$ and the colour
factor to $N_C = \num{3.2(13)}$.
All resuls are in good agreement to the standard model and can be considered as
one of its best confirmations.
}

@@

Due to inefficiencies in energy measurement it is possible that not all the
energy provided by the apparatus can finally be seen. Therefore there are large
uncertainties in jet energy measurements, whereas these uncertainties are
smaller for muonic events. This fact can also be used to distinguish these
events from each other and from background as well. For example if the $Z^0$
decays to two very unstable tauons, these tauons decay very fast. Then two
neutrinos are created, but they cannot be measured, so that their contribution
to the visible energy is missing. The visible part of the tau decays resembles
hadronic events, so it is important to be able to separate this background from
the hadronic signal.

@@

\subsection{Determination of the leptonic decay width}
Through the knowledge of the overall cross section it is possible to calculate
the leptonic decay width $Γ_l$. First of all we assume that the decay width is
the same for all leptons. We can use the following formula for determining it
by using the fit parameter for the muonic cross section:
\begin{equation}
σ_0^μ = \frac{12π}{M_Z^2}\frac{Γ_l^2}{Γ_Z^2}.
\label{eqn:leptonicwidth}
\end{equation}
This equation can easily be solved to determine the resulting leptonic decay
width $Γ_l$, because all the other variables are known.
\begin{equation}
Γ_l = (91 \pm 2 \pm 1)\si{\mega\electronvolt}
\label{eqn:leptonicwidthresult}
\end{equation}
In order to calculate the leptonic decay width, it is also possible to utilize
the following equation:
\begin{equation}
Γ_l = \frac{1}{3}(Γ_Z - 3Γ_{\mathrm{Had}} - 3Γ_{ν_l}).
\end{equation}
$Γ_{\mathrm{Had}}$ can be calculated in the same way as the leptonic decay
width in \eref{leptonicwidth}. $Γ_{ν_l}$ can be determinde by using the
following relation:
\begin{eqnarray}
Γ_{ν_e} &=& \frac{G_FM_Z^3}{12\sqrt{2}π}\\
&=& (165.71 \pm 0.10 \pm 0.09)\si{\mega\electronvolt}
\end{eqnarray}
Now the leptonic decay width $Γ_l$ can be evaluated.
\begin{equation}
Γ_l = (83 \pm 4 \pm 3) \si{\mega\electronvolt}
\end{equation}
This is in the same order of magnitude as the result shown in
\eref{leptonicwidthresult}. Although these two results do not agree in
one standard deviation, they do agree in two standard deviations.
Therefore it can be assumed that a more precise measurement in the future
will enable physicists to find a better agreement.

\subsection{Determination of the hadronic decay width}
\label{sec:hadrwidth}
As we now know the leptonic decay width, it is also possible to calculate the
hadronic decay width $Γ_{\mathrm{had}}$. There exists an anti-proportional relation
between these two variables, with only known constants as factors. The value for
the hadronic decay width evaluates to
\begin{equation}
Γ_{\mathrm{had}} = (1.86 \pm 0.01 \pm 0.02) \si{\giga\electronvolt}.
\end{equation}
There is also another relation which can be used to determine the value of this
decay width:
\begin{eqnarray}
Γ_{\mathrm{had}} &=& \Gamma_Z - 3\Gamma_{\nu_e} - 3\Gamma_e\\
&=& (1.86 \pm 0.05 \pm 0.03) \si{\giga\electronvolt}.
\end{eqnarray}
It is clearly visible that these two results are in perfect agreement.

@@

The achieved results by now offer furthermore an oportunity to obtain the
weak mixing angle using the formula
\begin{eqnarray}
Γ_l &=& \frac{G_F M_Z^3}{24 \sqrt{2}π}\left(1+[1-4|Q_l|\sin² Θ_W]²\right)\,.
\end{eqnarray}
As $Γ_l$ as well as $M_Z$ is known, the expression $\sin² Θ_W$ can be computed.
\begin{eqnarray}
\sin^2\Theta_W &=& \frac{1-\sqrt{\frac{\Gamma_e\cdot 24\sqrt{2}\pi}{G_FM_Z^3}-1}}{4}\\
\label{eqn:gammaquark}
\nonumber &=& 0.23 \pm 0.05 \pm 0.08
\end{eqnarray}
Thereby the weak mixing angle is determined, too.
\begin{eqnarray}
Θ_W &=& \arcsin(\sqrt{\sin^2\Theta_W})\\
\nonumber &=& 0.50 \pm 0.11 \pm 0.07\\
\nonumber &=& (28 \pm 6 \pm 4)\symbol{23}
\end{eqnarray}

@@

Finally the measurements allow also to calculate the colour factor $N_C$ using
$Γ_{\mathrm{had}}$. The hadronic decay width was determined in \ref{sec:hadrwidth}.

In reference to the standard model there is a way to compute $N_C$.
\begin{equation}
 Γ_{\mathrm{had}} = N_C \cdot K_{QCD}\cdot ( 2Γ_u + 3Γ_d )
 \label{eqn:gammahad2}
\end{equation}
The decay widths $Γ_u$ and $Γ_d$ can be calculated by the standard model, using the
value of the Weinberg angle as seen in \eref{gammaquark}. The charges of the up-quark was
set to $\frac{2}{3}$ and the charge of the down-quark to $-\frac{1}{3}$. Then the colour
factor can be calculated.
\begin{equation}
 N_C = 3.20 \pm 0.11 \pm 0.07
\end{equation}
This result confirms, that there are no new quarks to be expected with a mass
lower than some \SI{90}{\GeV}.

@@

\section{ Summary}
In order to sum up the results one has to say that this is a very precise
confirmation of the predictions of the standard model. The first exact
measurement of the weak mixing angle was done, which was one of the most important
tasks of this experiment.
